---
title: Sexual dimorphism in Chinstrap penguins"
author: "Benedikt Haarscheidt, Jaesub Kim, Ahmed Tijani Akinfalabi "
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
---
```{r echo=FALSE, include=FALSE}

library(palmerpenguins)
library(datasets)
library(MASS)
library(effsize)
library(DAAG)
library(palmerpenguins)
```

This is the mammals dataset containing body and brain weigth of 62 land species
```{r}
data(mammals)
head(mammals)


```
This is the possum datase containing measurments of possums in australia 
```{r}
data(possum)
head(possum)
```

## Data preparation

Subsetting the penguin species 
```{r}
head(penguins)
df=penguins
adelieset <- df[df$species == "Adelie", ]
adeliemale =adelieset[adelieset$sex == "male",]
adeliefemale = adelieset[adelieset$sex =="female",]

gentooset=df[df$species =="Gentoo",]
gentoomale = gentooset[gentooset$sex =="male",]
gentoofemale = gentooset[gentooset$sex =="female",]

chinstrapset=df[df$species =="Chinstrap",]
chinstrapmale=chinstrapset[chinstrapset$sex =="male",]
chinstrapfemale=chinstrapset[chinstrapset$sex =="female",]

head(chinstrapmale)
head(chinstrapfemale)


# Standard deviation of Chinstrap Male

male_std_dev_bill_Length <- sd(chinstrapmale$bill_length_mm)
male_std_dev_bill_Length
male_std_dev_bill_depth <- sd(chinstrapmale$bill_depth_mm)
male_std_dev_bill_depth
male_std_dev_flipper_Length <- sd(chinstrapmale$flipper_length_mm)
male_std_dev_flipper_Length
male_std_dev_body_mass <- sd(chinstrapmale$body_mass_g)
male_std_dev_body_mass

# Standard deviation of Chinstrap Female

female_std_dev_bill_Length <- sd(chinstrapmale$bill_length_mm)
female_std_dev_bill_Length
female_std_dev_bill_depth <- sd(chinstrapmale$bill_depth_mm)
female_std_dev_bill_depth
female_std_dev_flipper_Length <- sd(chinstrapmale$flipper_length_mm)
female_std_dev_flipper_Length
female_std_dev_body_mass <- sd(chinstrapmale$body_mass_g)
female_std_dev_body_mass

```


```{r}
sextable=table(penguins$sex,penguins$species)
sextable
adeliesex=sextable[,1]
gentoosex=sextable[,3]
chinstrapsex=sextable[,2]

```



## Data analysis
numerical data summary adelie:
```{r}
#Important to take arithmetic mean 
meanbodymassadelie=mean(adelieset$body_mass_g, na.rm=TRUE)
meanbodymassadeliemale=mean(adeliemale$body_mass_g, na.rm=TRUE)
meanbodymassadeliefemale=mean(adeliefemale$body_mass_g, na.rm=TRUE)
medianbodymassadeliemale=median(adeliemale$body_mass_g, na.rm=TRUE)
medianbodymassadeliefemale=median(adeliefemale$body_mass_g, na.rm=TRUE)

summaryadeliefemale=summary(adeliefemale[,3:6])
summaryadeliemale=summary(adeliemale[,3:6])


```

numerical data summary chinstrap bodymass
```{r}
meanbodymasschinstrap=mean(chinstrapset$body_mass_g, na.rm=TRUE)
meanbodymasschinstrapmale=mean(chinstrapmale$body_mass_g, na.rm=TRUE)
meanbodymasschinstrapfemale=mean(chinstrapfemale$body_mass_g, na.rm=TRUE)
medianbodymasschinstrapmale=median(chinstrapmale$body_mass_g, na.rm=TRUE)
medianbodymasschinstrapfemale=median(chinstrapfemale$body_mass_g, na.rm=TRUE)

summarychinstrapfemale=summary(chinstrapfemale[,3:6])
summarychinstrapmale=summary(chinstrapmale[,3:6])
```

numerical data summary chinstrap billlength 
```{r}
meanbilllengthchinstrap=mean(chinstrapset$bill_length_mm, na.rm=TRUE)
meanbilllengthchinstrapmale=mean(chinstrapmale$bill_length_mm, na.rm=TRUE)
meanbilllengthchinstrapfemale=mean(chinstrapfemale$bill_length_mm, na.rm=TRUE)
medianbilllengthchinstrapmale=median(chinstrapmale$bill_length_mm, na.rm=TRUE)
medianbilllengthchinstrapfemale=median(chinstrapfemale$bill_length_mm, na.rm=TRUE)

summarychinstrapfemale=summary(chinstrapfemale[,3:6])
summarychinstrapmale=summary(chinstrapmale[,3:6])
```

numerical data summary billdepth 
```{r}
meanbilldepthchinstrap=mean(chinstrapset$bill_depth_mm, na.rm=TRUE)
meanbilldepthchinstrapmale=mean(chinstrapmale$bill_depth_mm, na.rm=TRUE)
meanbilldepthchinstrapfemale=mean(chinstrapfemale$bill_depth_mm, na.rm=TRUE)
medianbilldepthchinstrapmale=median(chinstrapmale$bill_depth_mm, na.rm=TRUE)
medianbilldepthchinstrapfemale=median(chinstrapfemale$bill_depth_mm, na.rm=TRUE)

summarychinstrapfemale=summary(chinstrapfemale[,3:6])
summarychinstrapmale=summary(chinstrapmale[,3:6])
```

numerical data summary flipperlength 
```{r}
meanflipperlengthchinstrap=mean(chinstrapset$flipper_length_mm, na.rm=TRUE)
meanflipperlengthchinstrapmale=mean(chinstrapmale$flipper_length_mm, na.rm=TRUE)
meanflipperlengthchinstrapfemale=mean(chinstrapfemale$flipper_length_mm, na.rm=TRUE)
medianflipperlengthchinstrapmale=median(chinstrapmale$flipper_length_mm, na.rm=TRUE)
medianflipperlengthchinstrapfemale=median(chinstrapfemale$flipper_length_mm, na.rm=TRUE)

summarychinstrapfemale=summary(chinstrapfemale[,3:6])
summarychinstrapmale=summary(chinstrapmale[,3:6])
```
### Shapiro-wilke test
check if the data sets of the penguins satisfy the normal distribution, to use 
subsequent t-test 

```{r label = shapiro-wilke-test, echo=FALSE}
a_bill_length_female = shapiro.test(na.omit(adeliefemale$bill_length_mm))$p.value
a_bill_length_male = shapiro.test(na.omit(adeliemale$bill_length_mm))$p.value

a_bill_depth_female = shapiro.test(na.omit(adeliefemale$bill_depth_mm))$p.value
a_bill_depth_male_fit = adeliemale$bill_depth_mm[adeliemale$bill_depth_mm<21.2]
a_bill_depth_male = shapiro.test(na.omit(a_bill_depth_male_fit))$p.value

a_flipper_length_female = shapiro.test(na.omit(adeliefemale$flipper_length_mm))$p.value
a_flipper_length_male = shapiro.test(adeliemale$flipper_length_mm)$p.value

a_body_mass_female = shapiro.test(na.omit(adeliefemale$body_mass_g))$p.value
a_body_mass_male = shapiro.test(na.omit(adeliemale$body_mass_g))$p.value


adelishapiro = data.frame(a_bill_length_female,a_bill_length_male,a_bill_depth_female,a_bill_depth_male, a_flipper_length_female,a_flipper_length_male,a_body_mass_female,a_body_mass_male)

adelishapiro>0.05

c_bill_length_female_fit = chinstrapfemale$bill_length_mm[chinstrapfemale$bill_length_mm<58]
c_bill_length_female = shapiro.test(na.omit(c_bill_length_female_fit))$p.value
c_bill_length_male = shapiro.test(na.omit(chinstrapmale$bill_length_mm))$p.value

c_bill_depth_female = shapiro.test(na.omit(chinstrapfemale$bill_depth_mm))$p.value
c_bill_depth_male = shapiro.test(na.omit(chinstrapmale$bill_depth_mm))$p.value

c_flipper_length_female = shapiro.test(na.omit(chinstrapfemale$flipper_length_mm))$p.value
c_flipper_length_male = shapiro.test(na.omit(chinstrapmale$flipper_length_mm))$p.value

c_body_mass_female = shapiro.test(na.omit(chinstrapfemale$body_mass_g))$p.value
c_body_mass_male = shapiro.test(na.omit(chinstrapmale$body_mass_g))$p.value


chinshapiro = data.frame(c_bill_length_female,c_bill_length_male,c_bill_depth_female,c_bill_depth_male, c_flipper_length_female,c_flipper_length_male,c_body_mass_female,c_body_mass_male)

chinshapiro>0.05

g_bill_length_female = shapiro.test(na.omit(gentoofemale$bill_length_mm))$p.value
g_bill_length_male_fit = gentoomale$bill_length_mm[gentoomale$bill_length_mm<59]
g_bill_length_male = shapiro.test(na.omit(g_bill_length_male_fit))$p.value

g_bill_depth_female = shapiro.test(na.omit(gentoofemale$bill_depth_mm))$p.value
g_bill_depth_male = shapiro.test(na.omit(gentoomale$bill_depth_mm))$p.value

g_flipper_length_female = shapiro.test(na.omit(gentoofemale$flipper_length_mm))$p.value
g_flipper_length_male = shapiro.test(na.omit(gentoomale$flipper_length_mm))$p.value

g_body_mass_female = shapiro.test(na.omit(gentoofemale$body_mass_g))$p.value
g_body_mass_male = shapiro.test(na.omit(gentoomale$body_mass_g))$p.value


gentooshapiro = data.frame(g_bill_length_female,g_bill_length_male,g_bill_depth_female,g_bill_depth_male, g_flipper_length_female,g_flipper_length_male,g_body_mass_female,g_body_mass_male)

gentooshapiro>0.05
```
To satisfy a normal distribution, 1 outline in each data set bill_depth_adelie_male, bill_length_chinstrap_female and bill_length_gentoo_male were removed.

### T-test

to check if there is a statistical significant sexual dimorphism, the t-test was used.

<<<<<<< HEAD



>>>>>>> 70d8008c88ee23f3744a13be020f12c19e1733de
=======
```{r label=t-test, echo=FALSE,results='hide'}
t.test(c_bill_length_female_fit,chinstrapmale$bill_length_mm)
t.test(chinstrapfemale$bill_depth_mm,chinstrapmale$bill_depth_mm)
t.test(chinstrapfemale$flipper_length_mm,chinstrapmale$flipper_length_mm)
t.test(chinstrapfemale$body_mass_g,chinstrapmale$body_mass_g)
```

|categories|female|male|p-value|significant difference|
|--|--:|--:|--:|:---:|
|bill length in mm|46.23|51.09|1e-13|yes|
|bill depth in mm|17.59|19.25|7e-13|yes|
|flipper length in mm|191.7|199.9|3e-07|yes|
|body mass in g|3527|3939|3e-07|yes|
  


## Effecsize by cohens d
The function calculates the mean of body_mass_g for males and females separately.
It then calculates the difference between these two means.
Finally, it standardizes this difference by dividing it by the pooled standard deviation of the two groups.

Output Interpretation:
The resulting Cohen's d value tells you how many standard deviations the means of the two groups (male and female) differ.
It's a measure of the effect size of the sex difference on body mass. 
```{r}

cohen.d(chinstrapset$body_mass_g, chinstrapset$sex)
cohen.d(chinstrapset$bill_length_mm, chinstrapset$sex)
cohen.d(chinstrapset$bill_depth_mm, chinstrapset$sex)
cohen.d(chinstrapset$flipper_length_mm, chinstrapset$sex)
```

```{r}
cohen.w()
```





>>>>>>> 4151dfb0f4bd5f0036c90e8bf9773b5841e2da9d

## Plots:

### Stripcharts for Adelie 

Stripcharts for all the attributes

```{r}
stripchart(adelieset$body_mass_g~adelieset$sex, main="Massdistribution between the biological sexes",method="stack", col=c(2,4))
stripchart(adelieset$bill_length_mm~adelieset$sex, main="Massdistribution between the biological sexes",method="stack", col=c(2,4))
stripchart(adelieset$flipper_length_mm~adelieset$sex, main="Massdistribution between the biological sexes",method="stack", col=c(2,4))
stripchart(adelieset$bill_depth_mm~adelieset$sex, main="Massdistribution between the biological sexes",method="stack", col=c(2,4))

```


### Stripcharts for Chinstrap

### Histograms bodyweight

```{r echo=FALSE}

## Summary statistics
xy <- summary(adeliemale$body_mass_g)



# Create a histogram
hist(adeliemale$body_mass_g,
     main = "Histogram of Body Mass of Male Adelie Penguins",
     xlab = "Body Mass Distribution in gram",
     ylab = "Density",
     col = "cyan1",
     mar = c(2, 1, 2, 1),
     oma = c(2, 1, 2, 1),
     xlim=c(3000,5000),
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(adeliemale$body_mass_g, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanbodymassadeliemale, col = "black", lwd = 2, lty = 2)
abline(v = medianbodymassadeliemale, col = "blue", lwd = 2, lty = 2)
abline(v = meanbodymassadeliefemale, col = "black", lwd = 2, lty = 1)
abline(v = medianbodymassadeliefemale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean male", "Median male", "Mean female", "Median female"), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.8)


```



```{r echo=FALSE}
## Summary statistics
yx <- summary(adeliefemale$body_mass_g)



# Create a histogram
hist(adeliefemale$body_mass_g,
     main = "Histogram of Body Mass of Female Adelie Penguins",
     xlab = "Body Mass Distribution in gram",
     ylab = "Density",
     col = "lightpink",
     mar = c(2, 1, 2, 1),
     oma = c(2, 1, 2, 1),
     xlim = c(2500,4500),
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(adeliefemale$body_mass_g, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanbodymassadeliefemale, col = "black", lwd = 2, lty = 2)
abline(v = medianbodymassadeliefemale, col = "blue", lwd = 2, lty = 2)
abline(v = meanbodymassadeliemale, col = "black", lwd = 2, lty = 1)
abline(v = medianbodymassadeliemale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean female", "Median female", "Mean male", "Median male"), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.75)

```

## Bodymass Chinstrap

```{r echo=FALSE}
## Summary statistics
yx <- summary(chinstrapmale$body_mass_g)



# Create a histogram
hist(chinstrapmale$body_mass_g,
     main = "Histogram of Body Mass of Male Chinstrap Penguins",
     xlab = "Body Mass Distribution in gram",
     ylab = "Density",
     col = "cyan1",
     mar = c(2, 1, 2, 1),
     oma = c(2, 1, 2, 1),
     xlim = c(2500,5000),
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(chinstrapmale$body_mass_g, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanbodymasschinstrapmale, col = "black", lwd = 2, lty = 2)
abline(v = medianbodymasschinstrapmale, col = "blue", lwd = 2, lty = 2)
abline(v = meanbodymasschinstrapfemale, col = "black", lwd = 2, lty = 1)
abline(v = medianbodymasschinstrapfemale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean male", "Median male","Mean female", "Median female" ), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.75)

```

```{r echo=FALSE}
## Summary statistics
yx <- summary(chinstrapfemale$body_mass_g)



# Create a histogram
hist(chinstrapfemale$body_mass_g,
     main = "Histogram of Body Mass of Female Chinstrap Penguins",
     xlab = "Body Mass Distribution in gram",
     ylab = "Density",
     col = "lightpink",
     mar = c(2, 1, 2, 1),
     oma = c(2, 1, 2, 1),
     xlim = c(2500,5000),
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(chinstrapfemale$body_mass_g, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanbodymasschinstrapfemale, col = "black", lwd = 2, lty = 2)
abline(v = medianbodymasschinstrapmale, col = "blue", lwd = 2, lty = 2)
abline(v = meanbodymasschinstrapmale, col = "black", lwd = 2, lty = 1)
abline(v = medianbodymasschinstrapmale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean female", "Median female", "Mean male", "Median male"), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.75)

```


```{r echo=FALSE, fig.align='center'}
# Set up the plotting area to have 1 row and 2 columns
par(mfrow=c(1, 2), mar = c(5, 4, 4, 2) + 0.1, oma = c(2, 1, 2, 1))

## Summary statistics for male
yx_male <- summary(chinstrapmale$body_mass_g)

# Histogram for male chinstrap penguins
hist(chinstrapmale$body_mass_g,
     main = "",
     xlab = "Body Mass (g)",
     ylab = "Density",
     col = "cyan1",
     xlim = c(2500, 5000),
     freq = FALSE)

# Add density line for male
lines(density(chinstrapmale$body_mass_g, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines for male
abline(v = meanbodymasschinstrapmale, col = "black", lwd = 2, lty = 2)
abline(v = medianbodymasschinstrapmale, col = "blue", lwd = 2, lty = 2)

## Summary statistics for female
yx_female <- summary(chinstrapfemale$body_mass_g)

# Histogram for female chinstrap penguins
hist(chinstrapfemale$body_mass_g,
     main = "",
     xlab = "Body Mass (g)",
     ylab = "",
     col = "lightpink",
     xlim = c(2500, 5000),
     freq = FALSE)

# Add density line for female
lines(density(chinstrapfemale$body_mass_g, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines for female
abline(v = meanbodymasschinstrapfemale, col = "black", lwd = 2, lty = 2)
abline(v = medianbodymasschinstrapfemale, col = "blue", lwd = 2, lty = 2)
# Add a common main title and a legend
mtext("Bill Length of Male and Female Chinstrap Penguins", side = 3, line = 2, outer = TRUE)
legend("topright", inset=c(-0.1, 0), 
       legend = c("Density", "Mean", "Median"), 
       col = c("black", "black", "blue"), lty = c(1, 2, 2), lwd = 2, cex=0.75, bty="n", xpd=TRUE)

# Reset par to default
par(mfrow=c(1, 1), mar = c(5, 4, 4, 2) + 0.1, oma = c(2, 1, 2, 1))

# Since we cannot have two legends for the shared elements, we mention these in the caption.







### Histogramm Bill length Chinstrap

```








```{r echo=FALSE}




# Create a histogram
hist(chinstrapmale$bill_length_mm,
 
     main = "Histogram of Bill Length of Male Chinstrap Penguins",
     xlab = "Bill length Distribution in mm",
     ylab = "Density",
     col = "cyan1",
     mar = c(2, 1,2, 1),
     oma = c(2, 1, 2, 1),
     xlim=c(40,60),
     
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(chinstrapmale$bill_length_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanbilllengthchinstrapmale, col = "black", lwd = 2, lty = 2)
abline(v = medianbilllengthchinstrapmale, col = "blue", lwd = 2, lty = 2)
abline(v = meanbilllengthchinstrapfemale, col = "black", lwd = 2, lty = 1)
abline(v = medianbilllengthchinstrapfemale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean male", "Median male", "Mean female", "Median female"), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.8)


```


```{r echo=FALSE}




# Create a histogram
hist(chinstrapfemale$bill_length_mm,
     main = "Histogram of Bill Length of Female Chinstrap Penguins",
     xlab = "Bill length Distribution in mm",
     ylab = "Density",
     col = "lightpink",
     mar = c(2, 1, 2, 1),
     oma = c(2, 1, 2, 1),
     xlim = c(40,60),
     ylim= c(0.0,0.3),
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(chinstrapfemale$bill_length_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanbilllengthchinstrapfemale, col = "black", lwd = 2, lty = 2)
abline(v = medianbilllengthchinstrapmale, col = "blue", lwd = 2, lty = 2)
abline(v = meanbilllengthchinstrapmale, col = "black", lwd = 2, lty = 1)
abline(v = medianbilllengthchinstrapmale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean female", "Median female", "Mean male", "Median male"), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.75)

```

```{r echo=FALSE, fig.align='center'}
# Set up the plotting area to have 1 row and 2 columns
par(mfrow=c(1, 2), mar = c(4, 4, 2, 1), oma = c(2, 2, 2, 1))

# Histogram for male chinstrap penguins' bill length
hist(chinstrapmale$bill_length_mm,
     main = "",
     xlab = "Bill Length (mm)",
     ylab = "Density",
     col = "cyan1",
     xlim = c(40, 60),
     ylim = c(0.0, 0.3),
     freq = FALSE)

# Add density line for male
lines(density(chinstrapmale$bill_length_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines for male
abline(v = meanbilllengthchinstrapmale, col = "black", lwd = 2, lty = 2)
abline(v = medianbilllengthchinstrapmale, col = "blue", lwd = 2, lty = 2)

# Histogram for female chinstrap penguins' bill length
hist(chinstrapfemale$bill_length_mm,
     main = "",
     xlab = "Bill Length (mm)",
     ylab = "",
     col = "lightpink",
     xlim = c(40, 60),
     ylim = c(0.0, 0.3),
     freq = FALSE)

# Add density line for female
lines(density(chinstrapfemale$bill_length_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines for female
abline(v = meanbilllengthchinstrapfemale, col = "black", lwd = 2, lty = 2)
abline(v = medianbilllengthchinstrapfemale, col = "blue", lwd = 2, lty = 2)

# Add a common main title and a legend
mtext("Bill Length of Male and Female Chinstrap Penguins", side = 3, line = 2, outer = TRUE)
legend("topright", inset=c(-0.1, 0), 
       legend = c("Density", "Mean", "Median"), 
       col = c("black", "black", "blue"), lty = c(1, 2, 2), lwd = 2, cex=0.75, bty="n", xpd=TRUE)


# Reset par to default for further plots
par(mfrow=c(1, 1), mar = c(5, 4, 4, 2) + 0.1, oma = c(2, 1, 2, 1))




```


### Histogram bill depth 
```{r echo=FALSE}




# Create a histogram
hist(chinstrapmale$bill_depth_mm,
   
     main = "Histogram of Bill Depth of Male Chinstrap Penguins",
     xlab = "Bill depth Distribution in mm",
     ylab = "Density",
     col = "cyan1",
     mar = c(2, 1,2, 1),
     oma = c(2, 1, 2, 1),
     xlim=c(15,22),
     
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(chinstrapmale$bill_depth_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanbilldepthchinstrapmale, col = "black", lwd = 2, lty = 2)
abline(v = medianbilldepthchinstrapmale, col = "blue", lwd = 2, lty = 2)
abline(v = meanbilldepthchinstrapfemale, col = "black", lwd = 2, lty = 1)
abline(v = medianbilldepthchinstrapfemale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean male", "Median male", "Mean female", "Median female"), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.8)


```


```{r echo=FALSE}




# Create a histogram
hist(chinstrapfemale$bill_depth_mm,
     main = "Histogram of Bill depth of Female Chinstrap Penguins",
     xlab = "Bill depth Distribution in mm",
     ylab = "Density",
     col = "lightpink",
     mar = c(2, 1, 2, 1),
     oma = c(2, 1, 2, 1),
     xlim = c(15,22),
     ylim= c(0.0,0.6),
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(chinstrapfemale$bill_depth_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanbilldepthchinstrapfemale, col = "black", lwd = 2, lty = 2)
abline(v = medianbilldepthchinstrapmale, col = "blue", lwd = 2, lty = 2)
abline(v = meanbilldepthchinstrapmale, col = "black", lwd = 2, lty = 1)
abline(v = medianbilldepthchinstrapmale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean female", "Median female", "Mean male", "Median male"), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.75)

```
```{r echo=FALSE, fig.align='center'}
# Set up the plotting area to have 1 row and 2 columns
par(mfrow=c(1, 2), mar = c(4, 4, 2, 1), oma = c(2, 2, 2, 1))

# Histogram for male chinstrap penguins' bill depth
hist(chinstrapmale$bill_depth_mm,
     main = "",
     xlab = "Bill Depth (mm)",
     ylab = "Density",
     col = "cyan1",
     xlim = c(15, 22),
     ylim= c(0.0, 0.6),
     freq = FALSE)

# Add density line for male
lines(density(chinstrapmale$bill_depth_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines for male
abline(v = meanbilldepthchinstrapmale, col = "black", lwd = 2, lty = 2)
abline(v = medianbilldepthchinstrapmale, col = "blue", lwd = 2, lty = 2)

# Histogram for female chinstrap penguins' bill depth
hist(chinstrapfemale$bill_depth_mm,
     main = "",
     xlab = "Bill Depth (mm)",
     ylab = "",
     col = "lightpink",
     xlim = c(15, 22),
     ylim= c(0.0, 0.6),
     freq = FALSE)

# Add density line for female
lines(density(chinstrapfemale$bill_depth_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines for female
abline(v = meanbilldepthchinstrapfemale, col = "black", lwd = 2, lty = 2)
abline(v = medianbilldepthchinstrapfemale, col = "blue", lwd = 2, lty = 2)

# Add a common main title above the plots
mtext("Bill Depth of Male and Female Chinstrap Penguins", side = 3, line = 1, outer = TRUE)

# Since we have a shared title, we remove the individual main titles by setting `main = ""` in the `hist` functions.

# Add a shared legend below the plots
# Add legend
legend("topright", inset=c(-0.1, 0), 
       legend = c("Density", "Mean", "Median"), 
       col = c("black", "black", "blue"), lty = c(1, 2, 2), lwd = 2, cex=0.75, bty="n", xpd=TRUE)


# Reset par to default for further plots
par(mfrow=c(1, 1), mar = c(5, 4, 4, 2) + 0.1, oma = c(2, 1, 2, 1))


```


## Histogram flipperlength
```{r echo=FALSE}




# Create a histogram
hist(chinstrapmale$flipper_length_mm,
   
     main = "Histogram of Flipper Length of Male Chinstrap Penguins",
     xlab = "Flipper Length Distribution in mm",
     ylab = "Density",
     col = "cyan1",
     mar = c(2, 1,2, 1),
     oma = c(2, 1, 2, 1),
     xlim=c(170,220),
     
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(chinstrapmale$flipper_length_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanflipperlengthchinstrapmale, col = "black", lwd = 2, lty = 2)
abline(v = medianflipperlengthchinstrapmale, col = "blue", lwd = 2, lty = 2)
abline(v = meanflipperlengthchinstrapfemale, col = "black", lwd = 2, lty = 1)
abline(v = medianflipperlengthchinstrapfemale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean male", "Median male", "Mean female", "Median female"), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.8)


```


```{r echo=FALSE}




# Create a histogram
hist(chinstrapfemale$flipper_length_mm,
     main = "Histogram of Flipper length of Female Chinstrap Penguins",
     xlab = "Flipper Length Distribution in mm",
     ylab = "Density",
     col = "lightpink",
     mar = c(2, 1, 2, 1),
     oma = c(2, 1, 2, 1),
     xlim = c(170,220),
     #ylim= c(0.0,0.6),
     freq = FALSE, # Using density instead of frequency
      # Set x-axis limits to cover full data range
     )     # Do not plot x-axis yet


# Add density line
lines(density(chinstrapfemale$flipper_length_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines
abline(v = meanflipperlengthchinstrapfemale, col = "black", lwd = 2, lty = 2)
abline(v = medianflipperlengthchinstrapmale, col = "blue", lwd = 2, lty = 2)
abline(v = meanflipperlengthchinstrapmale, col = "black", lwd = 2, lty = 1)
abline(v = medianflipperlengthchinstrapmale, col = "blue", lwd = 2, lty = 1)

# Add legend
legend("topright", legend = c("Density", "Mean female", "Median female", "Mean male", "Median male"), 
       col = c("black", "black", "blue", "black", "blue"), lty = c(1, 2, 2, 1, 1), lwd = 2,cex=0.75)

```


```{r echo=FALSE, fig.align='center'}
# Set up the plotting area to have 1 row and 2 columns
par(mfrow=c(1, 2), mar = c(5, 4, 4, 2) + 0.1, oma = c(0, 0, 2, 0))

# Histogram for male chinstrap penguins' flipper length
hist(chinstrapmale$flipper_length_mm,
     main = "",
     xlab = "Flipper Length (mm)",
     ylab = "Density",
     col = "cyan1",
     xlim = c(170, 220),
     ylim= c(0, 0.08),
     freq = FALSE)

# Add density line for male
lines(density(chinstrapmale$flipper_length_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines for male
abline(v = meanflipperlengthchinstrapmale, col = "black", lwd = 2, lty = 2)
abline(v = medianflipperlengthchinstrapmale, col = "blue", lwd = 2, lty = 2)

# Histogram for female chinstrap penguins' flipper length
hist(chinstrapfemale$flipper_length_mm,
     main = "",
     xlab = "Flipper Length (mm)",
     ylab = "",
     col = "lightpink",
     xlim = c(170, 220),
     ylim= c(0, 0.08),
     freq = FALSE)

# Add density line for female
lines(density(chinstrapfemale$flipper_length_mm, na.rm = TRUE), col = "black", lwd = 2)

# Add mean and median lines for female
abline(v = meanflipperlengthchinstrapfemale, col = "black", lwd = 2, lty = 2)
abline(v = medianflipperlengthchinstrapfemale, col = "blue", lwd = 2, lty = 2)

# Add a common main title and a legend
mtext("Histogram of Flipper Length of Chinstrap Penguins", side = 3, line = 1, outer = TRUE)
legend("topright", inset=c(-0.1, 0), 
       legend = c("Density", "Mean", "Median"), 
       col = c("black", "black", "blue"), lty = c(1, 2, 2), lwd = 2, cex=0.75, bty="n", xpd=TRUE)

# Reset par to default for further plots
par(mfrow=c(1, 1), mar = c(5, 4, 4, 2) + 0.1, oma = c(2, 1, 2, 1))


```


# chinstrap

## Datasummary numerical
```{r}
meanbodychinstrapmale=mean(adelieset$body_mass_g, na.rm=TRUE)
meanbodychinstrapfemale=mean(adeliemale$body_mass_g, na.rm=TRUE)
meanbodymassadeliefemale=mean(adeliefemale$body_mass_g, na.rm=TRUE)
medianbodymassadeliemale=median(adeliemale$body_mass_g, na.rm=TRUE)
medianbodymassadeliefemale=median(adeliefemale$body_mass_g, na.rm=TRUE)

summaryadeliefemale=summary(adeliefemale[,3:6])
summaryadeliemale=summary(adeliemale[,3:6])

```

